---
title: "Summative Assessment RM&DA"
author: "Your Name"
date: today
format:
  html:
    toc: true
  pdf:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom) # For cleaner model outputs
library(datasets) # Load the datasets package
```

### Short Brief

-   This is an open-book exam. You may use any resources available to you (notes, books, online documentation).

-   You are expected to work individually. Collaboration is not permitted.

-   All code and answers must be clearly presented in this document.

-   Ensure this document renders successfully to both an HTML and a PDF file.

-   Show all your work, including the R code used to arrive at your answers.

# Section 1: Data Wrangling

This section uses the `dataset`, which contains data about various car models.

### Step 1 - Load the database

Load the dataset

```{r}
# Place your code here

```

Inspect the dataset for basic characteristics such as type of variables, dimensions of the dataset, etc.

```{r}
# Place your code here

```

### Step 2 - Data wrangling

-   Create a new variable based on an existing one

The `am` column in the `mtcars` dataset represents the transmission type (0 = automatic, 1 = manual). Create a new factor variable called `transmission_type` based on the am column to acheive the result below:

```{r}
#| echo: false
mtcars %>% 
  select(am) %>% 
  mutate(transmission_type = factor(am, labels = c("Automatic", "Manual"))) %>% 
  head()
```

-   Create transmission_type

```{r}
# Place your code here

```

### Step 3 - Data Summarisation

Calculate the mean mpg (miles per gallon) for cars with 4 cylinders (cyl) and for cars with 6 cylinders.

- Calculate mean `mpg` for 4 and 6 cylinder cars

```{r}
#| echo: false
mtcars %>% 
  filter(cyl %in% c(4, 6)) %>% 
  group_by(cyl) %>% 
  summarise(mean_mpg = mean(mpg))
  
```


```{r}
# Place your code here
```


Question 4 (\[Insert points here\])

Filter the data to include only observations where the car has a manual transmission and has more than 4 gears. How many observations meet these criteria?

manual_and_more_than_4_gears \<- mtcars %\>% filter(am == 1 & gear \> 4) nrow(manual_and_more_than_4_gears)

Your answer here

Section 2: Data Visualization This section continues to use the mtcars dataset.

Question 5 (\[Insert Points Here\])

Create a scatter plot showing the relationship between mpg and wt (weight). Add appropriate labels for the axes and a title for the plot.

# Create scatter plot of mpg vs wt

ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point() + labs(title = "Relationship between MPG and Weight", x = "Weight (1000 lbs)", y = "MPG")

*Your written answer for Question 5 goes here.*

Question 6 (\[Insert Points Here\])

Generate a boxplot to visualize the distribution of mpg across different cyl (number of cylinders) values.

# Create boxplot of mpg by cyl

ggplot(mtcars, aes(x = factor(cyl), y = mpg)) + geom_boxplot() + labs(title = "MPG Distribution by Number of Cylinders", x = "Number of Cylinders", y = "MPG")

*Your written answer for Question 6 goes here.*

Question 7 (\[Insert Points Here\])

Create a bar chart showing the count of cars for each number of cylinders (cyl).

# Create bar chart of cylinder counts

ggplot(mtcars, aes(x = factor(cyl))) + geom_bar() + labs(title = "Car Count by Number of Cylinders", x = "Number of Cylinders", y = "Count")

*Your written answer for Question 7 goes here.*

Section 3: Basic Statistics This section uses the mtcars dataset.

Question 8 (\[Insert Points Here\])

Perform an independent samples t-test to compare the mean mpg for cars with automatic and manual transmissions. State the null and alternative hypotheses, report the p-value, and interpret the results.

# Perform t-test

t.test(mpg \~ am, data = mtcars)

*Your written answer for Question 8 goes here.*

Question 9 (\[Insert Points Here\])

Calculate the Pearson correlation coefficient between mpg and hp (horsepower). Interpret the strength and direction of the correlation.

# Calculate correlation

cor(mtcars$mpg, mtcars$hp)

*Your written answer for Question 9 goes here.*

Question 10 (\[Insert Points Here\])

Conduct a chi-squared test to examine the relationship between cyl (number of cylinders) and am (transmission type). State the null and alternative hypotheses, report the chi-squared statistic and p-value, and interpret the results.

# Perform chi-squared test

table_cyl_am \<- table(mtcars$cyl, mtcars$am) chisq.test(table_cyl_am)

*Your written answer for Question 10 goes here.*

Question 11 (\[Insert Points Here\])

Fit a linear model to predict mpg using wt (weight) and hp (horsepower). Provide the R code, summarize the model, and interpret the coefficients for wt and hp.

# Fit linear model

linear_model_mpg \<- lm(mpg \~ wt + hp, data = mtcars) summary(linear_model_mpg)

*Your written answer for Question 11 goes here.*

Question 12 (\[Insert Points Here\])

Fit a logistic regression model to predict am (transmission type) using mpg and wt (weight). Provide the R code, summarize the model, and interpret the odds ratio for mpg.

# Fit logistic model

logistic_model_am \<- glm(am \~ mpg + wt, data = mtcars, family = "binomial") summary(logistic_model_am) exp(coef(logistic_model_am))

*Your written answer for Question 12 goes here.*

Section 4: Basic Introduction to Multivariate Analysis This section uses the iris dataset.

Question 13 (\[Insert Points Here\])

Briefly explain the purpose of Principal Component Analysis (PCA). In the context of the \`iris
